Research Experience

### 2010-03-01 doc/presentation/thesis_pres.tex
Merge branch 'master' of ssh://oldcub.ath.cx/home/micahc/brainid

\documentclass{beamer}

\mode<presentation>
{
  \usetheme{Darmstadt}
  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}
\usefonttheme[onlylarge]{structurebold}
\setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
\setbeamertemplate{navigation symbols}{}
\setbeamercovered{transparent}

@@

\title{A Crash Course in FMRI Imaging}

\subtitle{Or More than you ever wanted to know about FMRI}

@@

\pgfdeclareimage[height=0.5cm]{university-logo}{VT_marn_she_invent_cmyk}
\logo{\pgfuseimage{university-logo}}

@@

\section{About FMRI}
\subsection{FMRI Basics}
\begin{frame}{Pulse Sequence}
  % - A title should summarize the slide in an understandable fashion
  %   for anyone how does not follow everything on the slide itself.

  \begin{itemize}
    \item $T2*$ Weighted Image, acquired either line by line or 
         with spiral sequence.
    \item One or more more slices are captured in a single reptition time (TR).
    \item Echo Planar Pulse Sequence, slices acquired in less than a
        tenth of a second, depending on resolution, etc.
    \item A full volume can usually be acquired in 2 to 4 seconds.
  \end{itemize}
  \begin{figure}
  \includegraphics[scale=.23]{epi}
  \caption{
    \tiny
    \href{http://airto.hosted.ats.ucla.edu/BMCweb/BMC_BIOS/MarkCohen/Papers/EPI-fMRI.html}
    {http://airto.hosted.ats.ucla.edu/} }
  \end{figure}
\end{frame}

\subsection{The Bold Response}
\begin{frame}{Visualization of the BOLD Response}
\begin{figure}
\includegraphics[scale=.23]{model}
\caption{
    \tiny
    \cite{ISI:000189252300007}
}
\end{figure}
\end{frame}

\begin{frame}{Equations}
  \begin{itemize}
    \item Normalized Cerebral Blood Flow:
    $$\ddot{f}(t) = \epsilon u(t) - \dot{f}(t)/\tau_s - (f(t)/\tau_f - 1)$$
    \item Normalized Cerebral Blood Volume:
    $$\dot{v}(t) = (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) $$
    \item Normalized Deoxyhaemoglobin Content:
    $$\dot{q}(t) = \frac{1}{\tau_0}\left(\frac{f(t)(1- (1-E_0)^{1/f(t)})}{E_0} - 
            \frac{q(t)}{v(t)^{1-1/\alpha}}\right)$$
    \item Hemodynamic Response - BOLD Signal
    $$y(t) = V_0(a_1( 1 - Q(t)) - a_2(1 - V(t)))$$
  \end{itemize}
\end{frame}

\subsection{Practical Concerns}
\begin{frame}{Pros and Cons}
  \begin{itemize}
    \item Benefits
    \begin{itemize}
        \item In Comparison to EEG/MEG very good resolution.
        \item Unlike MEG or EEG Image reconstruction is a known/solved problem.
    \end{itemize}
    \item Drawbacks
    \begin{itemize}
      \item Measures the result of a long chain of activations
      \begin{itemize}
        \item Exact variables and parameters are unknown and are
            difficult to calculate.
        \item Significant Amount of Lag between activation
            and a measurable output - can be as much as 8 seconds.
      \end{itemize}
      \item Slow Temporal Resolution
      \item Extremely Low Signal-to-Noise Ratio 
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Drawbacks Continued}
   \begin{itemize}
      \item Non-Gaussian, non-stationary Noise, Caused by:
      \begin{itemize}
        \item Physiological Interference
        \item Interference from the scanner
        \item Other, as yet unknown, factors
        \item See \cite{ISI:000080539300008} and \cite{ISI:000246766700003}
      \end{itemize}
  \end{itemize}
\begin{figure}
    \includegraphics[scale=.15]{timeseries.png}
\end{figure}
\end{frame}

\section{Processing}
\subsection{Preprocessing}
\begin{frame}{Preprocessing}
\begin{itemize}
    \item Gaussian Filters especially when using the General Linear Model
    \item Removing Drift
    \begin{itemize}
        \item Extremely necessary, otherwise p value must be set prohibitively low. \cite{ISI:000246766700003}
        \item High Pass filters
        \item Various De-Trending methods 
        \begin{itemize}
            \item Polynomial fit
            \item Spline Fit
            \item Subtracting the Mean from the Timeseries
            \item Wavelets
        \end{itemize}
    \end{itemize}
     \item Detrended signals are often normalized using \%Difference.
    \item per-activation normalization
    \begin{itemize}
        \item Average signal for a few samples before activation \emph{should} begin.
    \end{itemize}
\end{itemize}
\end{frame}

\subsection{Linear Processing}
\begin{frame}{The General Linear Model}
\begin{itemize}
    \item Uses linear methods to extract useful information
    \item Can be effective for certain types of signal
    \item Used in many of the classical studies to determine regions
        associated with various tasks.
    \item Limited in dealing with complicated tasks.
    \item Utilizes Statistical Parameter Maps used to create clusters, then find 
        clusters not likely to have arisen randomly. (Random Field Theory)
    \item SPM is the standard tool for performing linear analysis.
\end{itemize}
\end{frame}

\subsection{Nonlinear Processing}
\begin{frame}{Nonlinear Analysis}
\begin{itemize}
    \item Use nonlinear state evolution models to determine activation.
    \item Removal of drift still necessary.
    \item Typically Rely on some sort of gradient decent technique to determine
        model parameters.
    \item Neural Networks
    \begin{itemize}
        \item Good at gathering disparate low SNR signals into meaningful information. 
        \item The method typically used by people studying visual cortex to "read" images from 
            a persons brain.
        \item Needs some method of removing irrelevant voxel data, to prevent over-fitting.
    \end{itemize}
    \item Particle Filters
    \begin{itemize}
        \item Based on real biological systems - recall the BOLD state equations
        \item Able to determine underlying state variables
        \item Tolerant of Noisy readings
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{BOLD Model}
\begin{itemize}
    \item The Balloon Model proposed by \cite{ISI:000073759600002} is the basic model.
    \item There are more complicated versions of the BOLD model:
    \begin{itemize}
        \item \cite{ISI:000240969200015} Reviews several existing models
        \item \cite{Buxton2004S220} Pioneered the Balloon model which was shown in the beginning.
        \item \cite{ISI:000234015300018} Adds interesting neural activation and a habituation model
        \item Some models loosen the link between CMRO2 (oxygen metabolism)
            and Cerebral Blood Flow - likely due to several papers that report
            such a decoupling.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Post Stimulus Undershoot Controversy}
\begin{itemize}
    \item The delinking of CMRO2 from CBF is postulated based on the "Post Stimulus Undershoot"
    \begin{itemize}
        \item In 2008 several papers claimed prove CBV and CBF both returned
            to normal before the BOLD signal, implying a decoupling of CBV from the BOLD signal.
        \item In November of 2008, \cite{ISI:000265938700001} went out of the way to refute such a de-coupling,
            claiming that other experiments were catching Arterial CBV rather than Venous CBV.
        \item In \cite{ISI:000240969200015} (2004), it was shown that merely adding volume viscosity to
            the BOLD equations was sufficient to account for much of the undershoot.
    \end{itemize}
\end{itemize}
\end{frame}

@@

\section{Conclusion}
\begin{frame}{The Future of FMRI}
\begin{itemize}
    \item Extremely Powerful tool.
    \item Better Models are increasing FMRI's flexibility.
    \item Combining FMRI with other methods of imaging will add 
        extensive knowledge to the understanding of how the human brain
        works.
    \item Much of the newer research is based around non-linear non-Gaussian
        models.
\end{itemize}
\end{frame}

% All of the following is optional and typically not needed. 
\appendix
\section{References}

>>>>

\documentclass{beamer}
\usepackage[orientation=landscape,size=custom,width=16,height=10,scale=0.5,debug]{beamerposter}

\mode<presentation>
{
  \usetheme{VT}
  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}

\usefonttheme[onlylarge]{structurebold}
\usecolortheme{crane}
\setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
\setbeamertemplate{navigation symbols}{}
\setbeamercovered{transparent}

@@

\title{BOLD Parameter Estimation using Sequential Monte Carlo Methods}

%\subtitle{}

@@

\pgfdeclareimage[width=1.5cm]{university-logo}{logo}
\logo{\pgfuseimage{university-logo}}

@@

\section{FMRI Review}
\subsection{The Bold Response}
\begin{frame}{The BOLD Response}
\begin{figure}
\includegraphics[scale=.23]{model}
\caption{
    \tiny
    \cite{ISI:000189252300007}
}
\end{figure}
\end{frame}

\begin{frame}{Purpose/Usefullness}
  \begin{itemize}
    \item blah
  \end{itemize}
\end{frame}

\begin{frame}{BOLD Signal Properties}
  \begin{itemize}
    \item ...
    \begin{itemize}
      \item Exact variables and parameters are unknown and are
          difficult to calculate.
      \item Significant Amount of Lag between activation
          and a measurable output - can be as much as 8 seconds.
    \end{itemize}
    \item Slow Temporal Resolution
    \item Noise characterized by brownian motion
    \item fft of mri signal, with and without DC
  \end{itemize}
\end{frame}

\begin{frame}{Preprocessing}
  \begin{itemize}
    \item High Pass Filter and Low Pass Filter
    \item Wavelet Detrending
    \item ...
    \item Spline
  \end{itemize}
\end{frame}

\section{Statistical Parametric Mapping}
\begin{frame}{Method}
  \begin{itemize}
    \item
  \end{itemize}
\end{frame}

\begin{frame}{Results}
  \begin{itemize}
    \item ...
  \end{itemize}
\end{frame}

\begin{frame}{Limitations}
  \begin{itemize}
    \item ...
  \end{itemize}
\end{frame}

\section{Nonlinear Regression}
\begin{frame}{Equations}
  \begin{itemize}
    \item Normalized Cerebral Blood Flow:
    $$\ddot{f}(t) = \epsilon u(t) - \dot{f}(t)/\tau_s - (f(t)/\tau_f - 1)$$
    \item Normalized Cerebral Blood Volume:
    $$\dot{v}(t) = (1/\tau_0)( f(t) - v(t) ^ {1/\alpha}) $$
    \item Normalized Deoxyhaemoglobin Content:
    $$\dot{q}(t) = \frac{1}{\tau_0}\left(\frac{f(t)(1- (1-E_0)^{1/f(t)})}{E_0} - 
            \frac{q(t)}{v(t)^{1-1/\alpha}}\right)$$
    \item Hemodynamic Response - BOLD Signal
    $$y(t) = V_0(a_1( 1 - Q(t)) - a_2(1 - V(t)))$$
  \end{itemize}
\end{frame}

\begin{frame}{Alternative Equations}
\begin{itemize}
    \item The Balloon Model proposed by \cite{ISI:000073759600002} is the basic model.
    \item There are more complicated versions of the BOLD model:
    \begin{itemize}
        \item \cite{ISI:000240969200015} Reviews several existing models
        \item \cite{Buxton2004S220} Pioneered the Balloon model which was shown in the beginning.
        \item \cite{ISI:000234015300018} Adds interesting neural activation and a habituation model
        \item Some models loosen the link between CMRO2 (oxygen metabolism)
            and Cerebral Blood Flow - likely due to several papers that report
            such a decoupling.
    \end{itemize}
\end{itemize}
\end{frame}

@@

\section{Parameter Identification}
\begin{frame}{Particle Filters}
\begin{itemize}
    \item Why Particle Filters
    \item Goal: $\epsilon$
    \item Entire brain
\end{itemize}
\end{frame}

\begin{frame}{Single Timeseries Results}
\end{frame}

\begin{frame}{Simulation}
\end{frame}

\begin{frame}{Simulation Results}
\end{frame}

\section{Conclusion}

### 2010-03-01 doc/thesis/thesis.tex
Merge branch 'master' of ssh://oldcub.ath.cx/home/micahc/brainid

\begin{document}

>>>>

\begin{document}
\section*{Introduction}
Functional Magnetic Resonance Imaging (FMRI) is a powerful tool in the analysis
neural activity. At the cost of rather limited temporal resolution, FMRI gives
high spatial for relatively cheap. Whereas other methods
of analyzing neural signals can be invasive or ill-posed, 
FMRI is relatively quick and easy, and its analysis straight forward. Because of
these benefits, FMRI has great potential for studying human cognition.
Unfortunately, FMRI does suffer from severe noise and, as previously
mentioned, low temporal resolution. Because of the extensive noise, 
and relative unpredictability of the absolute signal level, a large amount 
of conditioning is necessary to obtain "valid" activation. In
fact, typical activation detection does little to account for variations
in the shape of the signal, but instead is a direct T-test between the measured signal
and the expected shape of the signal. Because of these limitations,
multiple comparison tests are imperative (ref fish), and active regions
must be very large to be detected with any confidence. This paper
will introduce a more robust method of analyzing FMRI signals for 
activation, by means of estimating the BOLD parameters.

\section*{General Linear Model}

\section*{BOLD Model}

\subsection*{Preprocessing}
After FMRI data has been acquired it is always necessary to modify the
data in some way to make different runs comparable. Because FMRI signal
levels are essentially unit-less, at the very least it is necessary to convert
the data into \% difference from the baseline. Finding the baseline is the
first hurdle in dealing with FMRI. Various methods exist for doing this,
but they are all ad-hoc. It is common to treat each stimulus pulse independently,
ignoring the long decay of the BOLD signal. By treating each input separately,
the "base" for calculating the percent difference is merely the signal level 
before the stimulus is applied. The problem with this is that the BOLD signal
can be heavily delayed, and fall times can be as long as 10 or 15 seconds,
meaning that the assumption is clearly a false one. Another popular method 
of dealing with this problem, is to put all the time series through a high
pass filter. This will of course remove the DC component of the signal, and 
some amount of so called "drift". The problem with this method is that it is
not adaptive to the input. Huge variations in drift frequencies can exist 
in a single time-series. Thus, a single cutoff frequency could miss a significant
drift component, or it could remove \emph{actual} signal, if the cutoff frequency is
set too high.

\subsection*{
